\documentclass[10pt,leqno ]{article}
    %The document class defines the master templates, the structure of the document, 
    %and lays out the types of 
    %objects that can be manipulated for this type of document. 
    %the brackets contain basic options that will be applied globally (throughout
    %the document). Here, we specify a 10pt font, and when we number an equation, the 
    %number will be on the left.
    %The document class is a file ***.cls. You will probably never have to edit or create
    % a .cls file. There are many available on the internet for your use.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amsfonts}
\usepackage[shortlabels]{enumitem}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{times}
\usepackage{amsthm}
\usepackage{hyperref}
%\usepackage{homework}
    %packages control the ``style'' or look of the document. These come in the form of 
    %files ***.sty. The package ``homework'' above was created by me. The other packages
    %are very common for this type of document. You can google to learn more about what
    %they can do, and what options they give you. For example
\usepackage{textcomp}
\usepackage[margin=1.5in]{geometry}
    %the geometry package lets you customize the margins of your document.
    % and the 

\usepackage{forest}
  % For drawing the tree illustrating the sample space  
\usepackage{setspace}
    %package gives us the ability to set the line spacing.
\usepackage{moreenum}   % for greek letters in enum
\usepackage{xcolor}

\newtheorem{theorem}{Theorem}
\theoremstyle{definition} 
\newtheorem{problem}[theorem]{Problem}
    %these set up environments for listing things. The numbering is automatic.
\usepackage{float}
\restylefloat{table}

\newenvironment{solution}[1][Solution]{\begin{doublespace}\textbf{#1.}\quad }{\ \rule{0.5em}{0.5em}\end{doublespace}}
    %this is the environment for writing solutions. Doble spaced, with an end of proof
    %box at the end
    
\title{Discrete Math\\
CSCI 150, Fall 2020\\
Homework 5}
\author{Guy Matz \\
Hunter College}
    %above is the information that goes in the title. Notice the { and }. 
    %the double slashes \\ mean start a new line.


\begin{document} %this means end the preamble (stuff controling the styles above and
%start the content of the document. We can make adjustments as we go. For example,
%\maketitle
\begin{problem} An odd number is an integer that can be written as $2k + 1$ where $k \in Z$.  Is -1 an odd number?
\\\\
\Large
With $k=-1$, -1 can be written as $2k + 1$: $2(-1) + 1 = -1$.  So, yes, -1 is an odd number.

\end{problem}
\newpage

\begin{problem} What is the contrapositive of “If it is raining, then I have my umbrella”.
\\\\
\Large
If it is not raining, then I do not have my umbrella.
\end{problem}
\newpage

\begin{problem} Is the following True or False?
$$ \forall x, y \in  \mathbb{Z}.(x > y) \implies (x^2 > y^2)$$
\Large
False!!  Counter-example: x=-1, y=-2
\\\\
To make the statement true, it could read:
$$ \forall x, y \in  \mathbb{N}, (x > y) \implies (x^2 > y^2)$$
\end{problem}
\newpage

\begin{problem} Show that there exist two irrational numbers $x, y$ such that $xy$ is rational.
\\\\
\Large
Let $x,y \notin \mathbb{Q}$, and $x = y^{-1}$.  Then $xy = 1$, and 1 is rational
\end{problem}
\newpage

\begin{problem} It is a  fact that if $n^2$ is a multiple of 3, then n is
a multiple of 3. Use this fact to show by contradiction that $\sqrt{3}$  is irrational.
\\\\
\Large
If $\sqrt{3} \in \mathbb{Q}$ then $\sqrt{3} = \frac{a}{b}, a,b \in \mathbb{Z}$.  Then $3b^2 = a^2$.  So $a^2$ must be divisbible by 3, and therefore $a$ is divisible by 3.  So let $a = 3k$ where $k \in \mathbb{Z}$ and we have $3b^2 = (3k)^2$, or $3b^2 = 9k^2$.  But now we have a contradiction since $b$ must be also divisible by 3, and we see that $\frac{a}{b}$ must be reducible since they have 3 as a common factor.  
\end{problem}
\newpage

\begin{problem} Prove by contradiction that $17n + 2$ is odd $\implies n$ is odd.
\\\\
\Large
Towards a contradiction, assume $n$ is even.  Then, for some $k \in \mathbb{Z}$, $n = 2k$.  We can then write our original equation as $17(2k) + 2$, or $34k + 2$.  We know that an even number multiplied by any other number results in an even number.  We also know that adding two even numbers results in an even number.  However we are told that the original sum results in an odd number, so  we have a contradiction and $n$ must be odd.
\end{problem}
\newpage


\begin{problem} Show using a direct proof and logical operators the following set equality.
$$A - B^c = A \cap B$$
\\\\
\Large
$x \in (A - B^c) \implies x \in A \wedge \neg x \notin B$.  But this is equal to $x \in A \wedge x \in B$, which is the same as $A \cap B$, so the two are equal.
\end{problem}
\newpage


\begin{problem} Given a set $S \subset \{1, 3, 5, 7, 9, .\dots \}$ Show that
$$ |P(S)| =  \sum_{i \in S}^{}i \implies |S| \text{ is even}$$
Hint:  The sum of elements in $S$ can be written as $2k+|S|$.
\\\\
\Large
Contrapositive:
$$|S| \text{ is odd} \implies |P(S)| \neq \sum_{i \in S}^{}i$$
Since the sum $\sum_{i \in S}^{}i = 2k + |S|$ we can say the sum is necessarily odd when $|S|$ is odd, since $2k$ is even, and an even number plus an odd number is odd.  We also know that $|P(S)|$ is even since $|P(S)| = 2^n$ for $|S| = n$.  Clearly $\sum_{i \in S}^{}$ and $|P(S)|$ are not equal if one is odd (the former) and the other is even.  We have shown the contrapositive holds, and therefore the original statement is true.
\end{problem}
\newpage


\begin{problem} Show using the contrapositive that
\begin{center}
$n^2$ is a multiple of 3 $\implies n$ is a multiple of 3
\end{center}
Hint:  Any number $n$ can be written as either $3k$,  $3k+ 1$,  or $3k+ 2$ for some integer $k$ (in the last two cases $n$ is not a multiple of 3).
\\\\
\Large
Contrapositive:
\begin{center}
$n$ is not a multiple of 3 $\implies n^2$ is not a multiple of 3
\end{center}
Since any number $n$ can be written as either $3k$, $3k+1$ or $3k+2$ for some integer $k$, we will examine the three cases in which $n$ is not a multiple of 3.
\\\\
Case 1: $n = 3k+1$\\
if $n = 3k+1$ then $n^2 = 9k^2 + 6k + 1$.  Attempting to factor out a 3 from this leaves us with 1 as a remainder, so it is NOT divisible by 3.
\\\\
Case 2: $n = 3k+2$\\
if $n = 3k+2$ then $n^2 = 9k^2 + 12k + 4$.  Attempting to factor out a 3 from this leaves us with 1 as a remainder, so it is NOT divisible by 3.
\\\\
By showing the contrapositive, that $n$ is not a multiple of 3 $\implies n^2$ is not a multiple of 3, we have shown that $n^2$ is a multiple of 3 $\implies n$ is a multiple of 3.
\end{problem}
\newpage


\begin{problem} What is $(\sqrt{2}^{\sqrt{2})^{\sqrt{2}}}$?  Use your finding to show that there exist two irrational numbers $x$ and $y$ such that $x^y$ is rational.
\\\\
\Large
$(\sqrt{2}^{\sqrt{2})^{\sqrt{2}}} = \sqrt{2}^2 = 2$.
\\\\
So let $x = \sqrt{2}^{\sqrt{2}}$ and $y = \sqrt{2}$.  
We see that there exist two irrational numbers $x$ and $y$ such that $x^y$ is rational.
\end{problem}
\newpage


\begin{problem} Show  by  contradiction  that  the  following  tiles  cannot  be  put  together  to make a perfect square.  Hint:  use a coloring argument similar to the one we saw in class.
\\\\
\Large
There are 4 blocks of 4 squares each, totaling 16, so the square would need to be 4x4.  It we paint the 4x4 grid as a chessboard we have 8 black squares and 8 white squares.  We see that 3 of the 4 blocks would occupt 2 white spaces and 2 black spaces.  The problem is with the "stalagmite" block, second from the left.  That block covers 3 of one color and only 1 of the other, so we end up covering 9 squares of one color, and only 7 of the other, which means we are not able to cover the square.
\end{problem}
\newpage


\end{document}
% testing for push to multiple remotes
